---
title: "Random matrix community simulations"
author: "Taryn Waite"
date: "9/2/2020"
output: html_document
---

# Load libraries, source functions, and set directory
```{r directory}
library(deSolve)
library(Matrix)
library(knitr)

source("simulation_functions.R")

dir <- "/Users/tarynwaite/Desktop/Mutualism project/LogisticMutualism/RunsOut"
dir <- "/Volumes/GoogleDrive/My Drive/Projects/WaiteProjects/Network"

# where to save the runs
save_dir <- dir
```

# Parameters
This is where we set parameters (k and r) and specify which community sizes (numbers of species) and dispersal values (range for the interaction matrix values) we want to test. We also set the time sequence to use for the numerical solver.
```{r parameters and ICs}
# parameters
#n_spp_vec <- (c(2, 3, 5, 8, 13, 21, 34, 55, 89, 144, 233, 377, 610, 987))
n_spp_vec <- c(55)
#n_spp_vec <- c(987)
#disp_vec <- c(0.001) 
#disp_vec <- 10^seq(from = -3, to = -1, length.out = 14)
disp_vec <- 10^seq(from = -3, to = -(1/5), length.out = 5)
K <- 10
r <- 0.05
parms <- list(r, K)

# time sequence
#max_time <- 1000
max_time <- 500
time.seq <- seq(from = 0, to = max_time, by = 0.1)
length_time <- length(time.seq)

# cutoff for extinction
eps <- 0.01
```

# The Loop
This is where we run lots of simulations and keep track of the results. ***Note: check the save_dir (in first chunk) BEFORE running
```{r Loop}

iterations <- 3
iter_vec <- 1:iterations
length_n_spp_vec <- length(n_spp_vec)
length_disp_vec <- length(disp_vec)
spp_dist_iter_combs <- length_n_spp_vec*length_disp_vec*iterations
seed_mat <- array(data = 1:spp_dist_iter_combs, dim = c(iterations, length_disp_vec, length_n_spp_vec))


run_list <- list()
start_run <- proc.time()
for (i in 1:length_n_spp_vec) {
  for (j in 1:length_disp_vec) {
    for (k in 1:iterations) {
      
      
      #========== store initial info ==========
      
      
      # set the seed for creation of the interaction matrix (unique seed for each matrix)
      iter_seed <- seed_mat[k, j, i]
      set.seed(iter_seed)
      
      # define local variables for number of species, dispersion, iteration number
      n_spp <- n_spp_vec[i]
      disp_val <- disp_vec[j]
      iter <- k
      
      # column vector of initial states
      n0 <- matrix(data = 10, nrow = n_spp, ncol = 1)
      
      # make a data frame with the seed, dispersion, number of species, and iteration number
      pars <- c("seed", "dispersion", "n_spp", "iteration")
      vals <- c(iter_seed, disp_val, n_spp, iter)
      init_df <- data.frame(pars, vals)
      # this data frame is the first item in the output list
      run_list[[1]] <- init_df

      
      #========== normal ==========
      
      
      # make an interaction matrix (norm) that is guaranteed to have mutualism 
      int_mat_norm <- generate_coef_mat2(n_spp, distribution = "rnorm", disp = disp_val)
      # separate out mutualism and non-mutualism
      mut_mat <- get_mut_mat(int_mat_norm)
      not_mut_mat<- get_not_mut_mat(int_mat_norm)
      # the normal distribution interaction matrix is the second item in the output list
      run_list[[2]] <- int_mat_norm

      
      #========== numerator ==========
      
      
      # run ode solver for numerator model
      out <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      #outNum_norm <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              #rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Num_norm <- expGrowth(out)
      #exp_Num_norm <- expGrowth(outNum_norm)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = out, expGrowth = exp_Num_norm, n_spp = n_spp)
      #extPers <- sumExtPer(out = outNum_norm, expGrowth = exp_Num_norm, n_spp = n_spp)
      sum_Num_persist_norm <- extPers[[1]]
      sum_Num_extnct_norm <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[3]] <- NA
      #if(!exp_Num_norm) {
        #run_list[[3]] <- outNum_norm}

      
      #========== denominator ==========
      
      
      # # run ode solver for denominator model
      out <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      #outDem_norm <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              #rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Dem_norm <- expGrowth(out)
      #exp_Dem_norm <- expGrowth(outDem_norm)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = out, expGrowth = exp_Dem_norm, n_spp = n_spp)
      #extPers <- sumExtPer(out = outDem_norm, expGrowth = exp_Dem_norm, n_spp = n_spp)
      sum_Dem_persist_norm <- extPers[[1]]
      sum_Dem_extnct_norm <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[4]] <- NA
      #if(!exp_Dem_norm) {
        #run_list[[4]] <- outDem_norm}
      
      
      # store normal results in a vector
      values_Num_norm <- c("Num", "norm", exp_Num_norm, sum_Num_persist_norm, 
                           sum_Num_extnct_norm)
      values_Dem_norm <- c("Dem", "norm", exp_Dem_norm, sum_Dem_persist_norm,
                           sum_Dem_extnct_norm)
      
      
      #========== uniform ==========
      
      
      # get an interaction matrix using uniform distribution
      int_mat_unif <- generate_coef_mat2(n_spp, distribution = "runif", disp = disp_val)
      # separate out mutualism and non-mutualism
      mut_mat <- get_mut_mat(int_mat_unif)
      not_mut_mat<- get_not_mut_mat(int_mat_unif)
      diag(not_mut_mat) <- 0
      # item 6 is the interaction matrix for uniform distribution
      run_list[[5]] <- int_mat_unif
    
      
      #========== numerator ==========
      
      
      # run ode solver for numerator model
      out <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      #outNum_unif <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              #rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Num_unif <- expGrowth(out)
      #exp_Num_unif <- expGrowth(outNum_unif)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = out, expGrowth = exp_Num_unif, n_spp = n_spp)
      #extPers <- sumExtPer(out = outNum_unif, expGrowth = exp_Num_unif, n_spp = n_spp)
      sum_Num_persist_unif <- extPers[[1]]
      sum_Num_extnct_unif <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[6]] <- NA
      #if(!exp_Num_unif) {
        #run_list[[6]] <- outNum_unif}

      
      #========== denominator ==========
      
      
      # run ode solver for denominator model
      out <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      #outDem_unif <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              #rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Dem_unif <- expGrowth(out)
      #exp_Dem_unif <- expGrowth(outDem_unif)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = out, expGrowth = exp_Dem_unif, n_spp = n_spp)
      #extPers <- sumExtPer(out = outDem_unif, expGrowth = exp_Dem_unif, n_spp = n_spp)
      sum_Dem_persist_unif <- extPers[[1]]
      sum_Dem_extnct_unif <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[7]] <- NA
      #if(!exp_Dem_unif) {
        #run_list[[7]] <- outDem_unif}


      # store uniform results in a vector
      values_Num_unif <- c("Num", "unif", exp_Num_unif, sum_Num_persist_unif,
                           sum_Num_extnct_unif)
      values_Dem_unif <- c("Dem", "unif", exp_Dem_unif, sum_Dem_persist_unif,
                           sum_Dem_extnct_unif)

      
      #========== final results ==========
      
      
      # build a dataframe for all results
      out_all_df <- data.frame(rbind(values_Num_norm, values_Dem_norm, values_Num_unif, values_Dem_unif))
      # colnames(out_all_df) <- c("NumODem", "Distribution", "Exp_growth", "Steps",  "S_persist",
      #                           "S_extinct")
      colnames(out_all_df) <- c("NumODem", "Distribution", "Exp_growth", "S_persist",
                                "S_extinct")
      # save results dataframe to output list
      run_list[[8]] <- out_all_df
      
      # save all info to a file
      file_name <- paste0(n_spp, "spp_", sprintf(x = disp_val, fmt = "%.3f"), "disp_", iter, "iter")
      saveRDS(object = run_list, file = paste0(save_dir, "/", file_name, ".Rda"))
      
    } # end k iteraction
  }   # end j dispersion value
}     # end i community size
end_run <- proc.time()
end_run-start_run
```

# Data Compilation
Here, we compile the data from the runs (run_list[[8]], the dataframe with final results) into one dataframe
```{r data comp}
#dir <- save_dir
dir <- "/Users/tarynwaite/Desktop/Mutualism project/LogisticMutualism/testData"
runs <- list.files(dir)

rdf <- data.frame()
for (run in runs){
  dirSave <- paste0(dir, "/", run)
  #run_list <- readRDS(dirSave)
  load(dirSave)
  
  disp_vec <- rep(x = run_list[[1]]$vals[2], times = 4)
  n_spp_vec <- rep(x = run_list[[1]]$vals[3], times = 4)
  iter_vec <- rep(x = run_list[[1]]$vals[4], times = 4)
  
  run_list_temp <- cbind(run_list[[8]], disp_vec, n_spp_vec, iter_vec)
  #run_list_temp <- cbind(run_list[[10]], disp_vec, n_spp_vec, iter_vec)
  rdf <- rbind(rdf, run_list_temp)
}

# save data 
saveRDS(rdf, file = "outData.RData")
```


