---
title: "Network: Cleaned up Version"
output: html_notebook
---

# Load libraries and set directory
```{r directory}
library(deSolve)
library(Matrix)
library(knitr)
dir <- "/Users/cmmoore/Documents/WaiteMutCode/RunsOut/"
#dir <- "/Volumes/GoogleDrive/My Drive/Projects/WaiteProjects/Network"
# where to save the runs
save_dir <- dir
```

# Matrix functions
These are functions used to create and extract information from coefficient matrices. This is where the matrices that are used in the systems of ODEs are created.
```{r matrix functions}

# creates a random coefficient matrix with a given number of species
generate_coef_mat <- function(n_spp, distribution = "rnorm", disp = 1){
  if (distribution == "rnorm") {
    IM_vals <- rnorm(n = n_spp*n_spp, mean = 0, sd = disp)
  } else if (distribution == "runif") {
    IM_vals <- runif(n = n_spp*n_spp, min = -disp, max = disp)
  } else {
    ### gives an error message instead of a warning (so that it won't try to access IM) 
    stop("generate_coef_mat must have a distribution of 'rnorm' or 'runif'" )
  }
  IM <- matrix(data = IM_vals, nrow = n_spp, ncol = n_spp)
  return(IM)
}

# takes a coefficient matrix and returns a list with mutualism indices 
get_indices <- function(mat){
  
  pos_ind <- which(mat > 0)
  pos_t_ind <- which(t(mat) > 0) 
  
  # index values that are positive in both original matrix and transpose
  mut_ind <- pos_ind[which((pos_ind) %in% (pos_t_ind))]
  
  return(mut_ind)
}

# takes a coefficient matrix and creates a matrix with only mutualisms
get_mut_mat <- function(mat){
  # get mutualism indices
  mut_ind <- get_indices(mat)
  
  # Create a blank matrix and fill with only mutualisms
  mut_mat <- matrix(data = 0, nrow = nrow(mat), ncol = nrow(mat))
  mut_mat[mut_ind] <- mat[mut_ind]
  
  # remove the values on the diagonal (in the previous steps these are counted as mutualisms)
  diag(mut_mat) <- 0
  
  return (mut_mat)
}

# generates a coefficient matrix that contains at least one mutualistic interaction
generate_coef_mat2 <- function(n_spp, distribution = "rnorm", disp = 1){
  # generate a random coefficient matrix
  mat <- generate_coef_mat(n_spp, distribution, disp)
  mut_mat <- get_mut_mat(mat)
  n_mut <- nnzero(mut_mat)
  while(n_mut == 0){
    mat <- generate_coef_mat(n_spp, distribution, disp)
    mut_mat <- get_mut_mat(mat)
    n_mut <- nnzero(mut_mat)
  }
  
  diag(mat) <- 0
  return(mat)
}

# takes a coefficient matrix and returns a matrix with only non-mutualism interactions
get_not_mut_mat <- function(mat){
  # get indices for mutualism
  mut_ind <- get_indices(mat)
  
  # index values not in the mutualism matrix
  all_ind <- c(1:nrow(mat)^2)
  not_mut_ind <- all_ind[which(!(all_ind %in% mut_ind))]
  
  # Create a blank matrix and fill with only non-mutualistic interactions
  not_mut_mat <- matrix(data = 0, nrow = nrow(mat), ncol = nrow(mat))
  not_mut_mat[not_mut_ind] <- mat[not_mut_ind]
  
  # make diagonals 0 (don't want intraspecific effects since we're using K)
  diag(not_mut_mat) <- 0 
  
  return(not_mut_mat)
}
```

# Other functions
These are other functions used in the loop.
```{r other functions}
# given desolve output, find out how many species persisted
get_num_persist <- function(out){
  num_persist <- 0
      for (sp in out[nrow(out), 2:ncol(out)]){
        if(!is.na(sp)){
          if (sp > 0){
          num_persist <- num_persist + 1
        }}}
  return(num_persist)
}

# tryCatch error handler for exponential growth
errorFunc <- function(e){
  unbounded <- integer(1)
  return(unbounded)
}

# run ode solver with tryCatch, return deSolve object if no exp growth, return integer(1) if exp growth
odeSolve <- function (y, times, func, parms, method, rootfun, events){
  result <- tryCatch({
              ode(y = y, times = times, func = func, parms = parms, method = method,
                  rootfun = rootfun, events = events)
            }, error = errorFunc)
  return(result)
}

# return true if there was exponential growth
expGrowth <- function(out){
  if(is.integer(out)){
    return(TRUE)
  }
  return(FALSE)
}

# count up extinctions and persistances (NA if there was exp growth)
#returns a list of persistances and extinctions
sumExtPer <- function(out, expGrowth, n_spp){
  if (expGrowth){
    extinct <- NA
    persist <- NA
  } else{
      persist <- get_num_persist(out)
      extinct <- n_spp - persist
  }
  return(list(persist, extinct))
}
```


# Parameters
This is where we set parameters (k and r) and specify which community sizes (numbers of species) and dispersal values (range for the interaction matrix values) we want to test. We also set the time sequence to use for the numerical solver.
```{r parameters and ICs}
# parameters
n_spp_vec <- (c(2, 3, 5, 8, 13, 21, 34, 55, 89, 144, 233, 377, 610, 987))
#n_spp_vec <- c(2, 3, 5, 8, 13, 21, 34, 55, 89, 144)
#n_spp_vec <- c(987)
#disp_vec <- c(0.001) 
#disp_vec <- 10^seq(from = -3, to = -1, length.out = 14)
disp_vec <- 10^seq(from = -3, to = -(1/5), length.out = 10)
K <- 10
r <- 0.05
parms <- list(r, K)

# time sequence
max_time <- 1000
time.seq <- seq(from = 0, to = max_time, by = 0.1)
length_time <- length(time.seq)
```

# Mutualism models
These are the two systems of differential equations that model population growth in a community.  
The ODE solver uses these systems to numerically solve.
```{r models}
NumeratorMut <- function(t, n, parms){
   with(as.list(c(n, parms)), {
     dns <- r*n*((k - n + mut_mat%*%n + not_mut_mat%*%n)/k)
    return(list(c(dns)))
  })
}

DenominatorMut <- function(t, n, parms){
  with(as.list(c(n, parms)), {
    dns <- r*n*((k - n + mut_mat%*%n + not_mut_mat%*%n)/(k + mut_mat%*%n))
    return(list(c(dns)))
  })
}

```

# Events
Here, we create the events that tell the ODE solver how to handle extinction and exponential growth.  
Using these events, the solver will set a species density equal to 0 if it is headed towards extinction (<0.001) and it will stop solving and throw an error if any species is headed towards exponential growth (>1000)
```{r events}
eps <- 0.01
## event triggered if state variable <= 0.01 (extinction) or if state variable >= 1000 (unbounded growth)
rootfun <- function (t, n, parms) {
  n1 <- rep(NA, length(n))
  for(i in seq(length(n))){
    if( n[i] > 10){
      n1[i] <- n[i]-1000
    }
    else{
      n1[i] <- n[i] - eps
    }
  }
  return(as.vector(n1))
}


## sets state variable = 0 if extinct or stops executing and produces an error if unbounded growth is happening
eventfun <- function(t, n, parms) {
  for (i in seq(length(n))){
    if (n[i] <= eps){
     n[i] <-0 
    } 
    else if (n[i] >= 1000){
      stop("error: unbounded growth")
    }
  }
  return(as.vector(n))
}
```


# The Loop
This is where we run lots of simulations and keep track of the results. ***Note: check the save_dir (in first chunk) BEFORE running
```{r Loop}

iterations <- 5
iter_vec <- 1:iterations
length_n_spp_vec <- length(n_spp_vec)
length_disp_vec <- length(disp_vec)
spp_dist_iter_combs <- length_n_spp_vec*length_disp_vec*iterations
seed_mat <- array(data = 1:spp_dist_iter_combs, dim = c(iterations, length_disp_vec, length_n_spp_vec))


run_list <- list()
start_run <- proc.time()
for (i in 1:length_n_spp_vec) {
  for (j in 1:length_disp_vec) {
    for (k in 1:iterations) {
      
      
      ###################################### store initial info #############################################
      
      
      # set the seed for creation of the interaction matrix (unique seed for each matrix)
      iter_seed <- seed_mat[k, j, i]
      #set.seed(iter_seed)
      
      # define local variables for number of species, dispersion, iteration number
      n_spp <- n_spp_vec[i]
      disp_val <- disp_vec[j]
      iter <- k
      
      # column vector of initial states
      n0 <- matrix(data = 10, nrow = n_spp, ncol = 1)
      
      # make a data frame with the seed, dispersion, number of species, and iteration number
      pars <- c("seed", "dispersion", "n_spp", "iteration")
      vals <- c(iter_seed, disp_val, n_spp, iter)
      init_df <- data.frame(pars, vals)
      # this data frame is the first item in the output list
      run_list[[1]] <- init_df

      
      ############################################# normal ###################################################
      
      
      # make an interaction matrix (norm) that is guaranteed to have mutualism 
      int_mat_norm <- generate_coef_mat2(n_spp, distribution = "rnorm", disp = disp_val)
      # separate out mutualism and non-mutualism
      mut_mat <- get_mut_mat(int_mat_norm)
      not_mut_mat<- get_not_mut_mat(int_mat_norm)
      # the normal distribution interaction matrix is the second item in the output list
      run_list[[2]] <- int_mat_norm

      
      ########################################### numerator #################################################
      
      
      # run ode solver for numerator model
      outNum_norm <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Num_norm <- expGrowth(outNum_norm)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = outNum_norm, expGrowth = exp_Num_norm, n_spp = n_spp)
      sum_Num_persist_norm <- extPers[[1]]
      sum_Num_extnct_norm <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[3]] <- NA
      if(!exp_Num_norm) {
        run_list[[3]] <- outNum_norm}

      
      ########################################### denominator #################################################
      
      
      # # run ode solver for denominator model
      outDem_norm <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Dem_norm <- expGrowth(outDem_norm)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = outDem_norm, expGrowth = exp_Dem_norm, n_spp = n_spp)
      sum_Dem_persist_norm <- extPers[[1]]
      sum_Dem_extnct_norm <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[4]] <- NA
      if(!exp_Dem_norm) {
        run_list[[4]] <- outDem_norm}
      
      
      # store normal results in a vector
      values_Num_norm <- c("Num", "norm", exp_Num_norm, sum_Num_persist_norm, 
                           sum_Num_extnct_norm)
      values_Dem_norm <- c("Dem", "norm", exp_Dem_norm, sum_Dem_persist_norm,
                           sum_Dem_extnct_norm)
      
      
      ########################################### uniform ##################################################
      
      
      # get an interaction matrix using uniform distribution
      int_mat_unif <- generate_coef_mat2(n_spp, distribution = "runif", disp = disp_val)
      # separate out mutualism and non-mutualism
      mut_mat <- get_mut_mat(int_mat_unif)
      not_mut_mat<- get_not_mut_mat(int_mat_unif)
      diag(not_mut_mat) <- 0
      # item 6 is the interaction matrix for uniform distribution
      run_list[[5]] <- int_mat_unif
    
      
      ########################################### numerator #################################################
      
      
      # run ode solver for numerator model
      outNum_unif <- odeSolve(y = n0, times = time.seq, func = NumeratorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Num_unif <- expGrowth(outNum_unif)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = outNum_unif, expGrowth = exp_Num_unif, n_spp = n_spp)
      sum_Num_persist_unif <- extPers[[1]]
      sum_Num_extnct_unif <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[6]] <- NA
      if(!exp_Num_unif) {
        run_list[[6]] <- outNum_unif}

      
      ########################################### denominator #################################################
      
      
      # run ode solver for denominator model
      outDem_unif <- odeSolve(y = n0, times = time.seq, func = DenominatorMut, parms = parms, method = "lsoda",
                              rootfun = rootfun, events = list(func = eventfun, root = TRUE))
      # check for exponential growth
      exp_Dem_unif <- expGrowth(outDem_unif)
      # count up extinctions and persistances
      extPers <- sumExtPer(out = outDem_unif, expGrowth = exp_Dem_unif, n_spp = n_spp)
      sum_Dem_persist_unif <- extPers[[1]]
      sum_Dem_extnct_unif <- extPers[[2]]
      # add time series to output list (NA if unbounded growth)
      run_list[[7]] <- NA
      if(!exp_Dem_unif) {
        run_list[[7]] <- outDem_unif}


      # store uniform results in a vector
      values_Num_unif <- c("Num", "unif", exp_Num_unif, sum_Num_persist_unif,
                           sum_Num_extnct_unif)
      values_Dem_unif <- c("Dem", "unif", exp_Dem_unif, sum_Dem_persist_unif,
                           sum_Dem_extnct_unif)

      
      ########################################## final results #############################################
      
      
      # build a dataframe for all results
      out_all_df <- data.frame(rbind(values_Num_norm, values_Dem_norm, values_Num_unif, values_Dem_unif))
      # colnames(out_all_df) <- c("NumODem", "Distribution", "Exp_growth", "Steps",  "S_persist",
      #                           "S_extinct")
      colnames(out_all_df) <- c("NumODem", "Distribution", "Exp_growth", "S_persist",
                                "S_extinct")
      # save results dataframe to output list
      run_list[[8]] <- out_all_df
      
      # save all info to a file
      file_name <- paste0(sprintf(x = n_spp, fmt = "%03d"), "spp_", sprintf(x = abs(log(x = disp_val, base = 10)), fmt = "%.2f"), "disp_", sprintf(x = iter, fmt = "%03d"), "iter")
      save(x = run_list, file = paste0(save_dir, file_name, ".Rda"))
      
    } # end k iteraction
  }   # end j dispersion value
}     # end i community size
end_run <- proc.time()
end_run-start_run
```

# Data Compilation
Here, we compile the data from the runs (run_list[[8]], the dataframe with final results) into one dataframe
```{r data comp}
dir <- save_dir
runs <- list.files(dir)

rdf <- data.frame()
for (run in runs){
  dirSave <- paste0(dir, "/", run)
  load(dirSave)
  
  disp_vec <- rep(x = run_list[[1]]$vals[2], times = 4)
  n_spp_vec <- rep(x = run_list[[1]]$vals[3], times = 4)
  iter_vec <- rep(x = run_list[[1]]$vals[4], times = 4)
  
  run_list_temp <- cbind(run_list[[8]], disp_vec, n_spp_vec, iter_vec)
  rdf <- rbind(rdf, run_list_temp)
}

```
