---
title: "Simulation Data Analysis"
author: "Taryn Waite"
date: "9/2/2020"
output: html_document
---

```{r setup, include=FALSE}
# read in the output data from the simulation
data <- readRDS("outData.RData")

library(tidyverse)
library(colorspace)
```

Creating some modified dataframes with information about persistance and exponential growth
```{r data wrangling}
# convert some column types and add a proportion persist column
data2 <- data %>% 
  mutate(S_persist = as.numeric(paste(S_persist)), S_extinct = as.numeric(paste(S_extinct)),
         prop_persist = as.numeric(paste(S_persist))/n_spp_vec,
         Exp_growth = as.character(Exp_growth) == "TRUE")

# persistance data grouped by model type, #species, dispersion
persistance_data <- data2 %>% 
  drop_na() %>% 
  group_by(NumODem, n_spp_vec, disp_vec) %>% 
  summarize(mean_prop_persist = mean(prop_persist, na.omit = T))

# with distributions separated 
persistance_data2 <- data2 %>% 
  drop_na() %>% 
  group_by(NumODem, n_spp_vec, disp_vec, Distribution) %>% 
  summarize(mean_prop_persist = mean(prop_persist, na.omit = T))

persistance_spread <- data2 %>% 
  drop_na() %>% 
  group_by(NumODem, n_spp_vec, disp_vec) %>% 
  summarize(mean_prop_persist = mean(prop_persist, na.omit = T)) %>% 
  tidyr::spread(key = NumODem, value = mean_prop_persist) %>% 
  rename(num_prop_persist = Num, dem_prop_persist = Dem)

# with distributions separated
persistance_spread2 <- data2 %>% 
  drop_na() %>% 
  group_by(NumODem, n_spp_vec, disp_vec, Distribution) %>% 
  summarize(mean_prop_persist = mean(prop_persist, na.omit = T)) %>% 
  tidyr::spread(key = NumODem, value = mean_prop_persist) %>% 
  rename(num_prop_persist = Num, dem_prop_persist = Dem)


# exp growth data grouped by model type, #species, dispersion
exp_growth_data <- data2 %>% 
  group_by(NumODem, n_spp_vec, disp_vec) %>% 
  summarize(prop_exp_growth = sum(Exp_growth)/n())

exp_growth_data2 <- data2 %>% 
  group_by(NumODem, n_spp_vec, disp_vec, Distribution) %>% 
  summarize(prop_exp_growth = sum(Exp_growth)/n())

exp_growth_spread <- data2 %>% 
  group_by(NumODem, n_spp_vec, disp_vec) %>% 
  summarize(prop_exp_growth = sum(Exp_growth)/n()) %>% 
  tidyr::spread(key = NumODem, value = prop_exp_growth) %>% 
  rename(num_prop_exp = Num, dem_prop_exp = Dem)

# with distributions separated
exp_growth_spread2 <- data2 %>% 
  group_by(NumODem, n_spp_vec, disp_vec, Distribution) %>% 
  summarize(prop_exp_growth = sum(Exp_growth)/n()) %>% 
  tidyr::spread(key = NumODem, value = prop_exp_growth) %>% 
  rename(num_prop_exp = Num, dem_prop_exp = Dem)


```

Heatmaps showing proportion persisted by number of species and dispersion (one for each model type) -- note, runs with exponential growth cannot be included (these are the blank regions in the upper right)
```{r proportion persisted}
# heatmaps of proportion persisted by number of species and dispersion
persistance_data %>% 
  filter(NumODem == "Num") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), fill = mean_prop_persist)) +
  geom_tile()
  
persistance_data %>% 
  filter(NumODem == "Dem") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), fill = mean_prop_persist)) +
  geom_tile()

# difference between models
persistance_spread %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_persist - dem_prop_persist)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Species Persisting \n(Numerator - Denominator)")

# difference between models, just uniform
persistance_spread2 %>%
  filter(Distribution == "unif") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_persist - dem_prop_persist)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Species Persisting \n(Numerator - Denominator)")

# difference between models, just normal
persistance_spread2 %>%
  filter(Distribution == "norm") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_persist - dem_prop_persist)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Species Persisting \n(Numerator - Denominator)")

```
Heatmaps showing proportion exponential growth by number of species and dispersion (one for each model type)
```{r proportion exponential growth}
# heatmaps of proportion exp.growth by number of species and dispersion
exp_growth_data %>% 
  filter(NumODem == "Num") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), fill = prop_exp_growth)) +
  geom_tile()
  
exp_growth_data %>% 
  filter(NumODem == "Dem") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), fill = prop_exp_growth)) +
  geom_tile()

# difference between models
exp_growth_spread %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_exp - dem_prop_exp)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Exponential Growth \n(Numerator - Denominator)")

# just uniform
exp_growth_spread2 %>% 
  filter(Distribution == "unif") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_exp - dem_prop_exp)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Exponential Growth \n(Numerator - Denominator)")

# just normal
exp_growth_spread2 %>% 
  filter(Distribution == "norm") %>% 
  ggplot(aes(x = as.factor(n_spp_vec), y = as.factor(disp_vec), 
             fill = num_prop_exp - dem_prop_exp)) +
  geom_tile() +
  scale_fill_continuous_divergingx(palette = 'RdBu', mid = 0, 
                                   name = "Difference in \nProportion Exponential Growth \n(Numerator - Denominator)")
```

```{r}

```

